MT @HanBaldwin: Fightin OOVs in German \#twitter

### 2013-07-27 darmstadt_gscl_2013/article_twitter_normalization.tex
slight update of article

  This article gives an overview of existing approaches to the problem
  of out-of-vocabulary (OOV) tokens and noisiness phenomena in natural
  language texts. These approaches are classified with regard to the
  size of text spans and knowledge inference mechanisms which they use
  in their work. Subsequently, quantitative and qualitative analyses
  of unknown words are conducted on German Twitter messages, in order
  to see how relevant the OOV issue is for this particular kind of
  micro-texts and to determine which peculiarities are especially
  characteristic for them. In a concluding step, we present a set of
  ad-hoc techniques which are supposed to tackle some of the most
  prominent disturbing effects found during the analyses and show how
  this set of techniques helps us to lower the average rate of
  out-of-vocabulary tokens in Twitter messages thereby improving their
  automatic Part-of-speech tagging quality.

@@

Though the problem of out-of-vocabulary (OOV) words and its closely
related task of textual normalization have been extensively studied in
computational linguistics since as early as the late 1950s
(cf. Petersen, \cite{petersen}) and were certainly anything but new at
the time when mobile communication emerged, it were small messages
that revived interest in this research area in the past two
decades.

Starting from the second half of 1990-ies, the rise of SMS messages,
growing popularity of online chats and launch of social networks lead
to a significant boost in the number of publications on preprocessing
casually typed texts. But despite this increased number of researches,
the main focus of the majority of these works was on English data with
a few exceptions of works on Spanish (Oliva et al., \cite{oliva}) and
French (Beaufort et al., \cite{beaufort}).

%% One of the first scientific milestones which marked the renaissance of
%% OOV studies in CL was a comprehensive article by Karen Kukich
%% published in the renowned ACM journal on December 4-th 1992
%% \cite{kukich}. By an interesting coincidence, just exactly one day
%% before that, a British engineer called Neil Papworth had sent the
%% world's first ever SMS-message \cite{guardian:sms}. But in contrast to
%% Papworth's SMS which only said ``Merry Christmas'' to one of his
%% friends, Kukich's work comprised more than 60 pages and provided a
%% fully-fledged review of the state of the art techniques for dealing
%% with unknown words within the scope of spelling correction programs.

In the next Section we will give a short overview of recent scientific
approaches to the problem of out-of-vocabulary words in non-standard
texts. After that, in Section \ref{error:analysis} we will analyze
which types of noisiness phenomena are especially characteristic for
German Twitter. Section \ref{normalization} will subsequently describe
an automatic procedure for mitigating some of the most obvious of
those effects. In a concluding step, we will perform an evaluation of
the results of this procedure and give further suggestions for future
research.

@@

Before proceeding with the description of existing methods for noisy
text normalization (NTN), we first would like to define the criteria
by which these methods could be classified. It should be noted that
there already exist several classifications of NTN methods including,
for example, Kukich (\cite{kukich}), Kobus et al. (\cite{kobus}), and
Sproat et al. (\cite{sproat}). But all of these classification
attempts seem to have one common shortcoming, namely they try to
involve several independent criteria within one classification scheme.

@@

Kobus (\cite{kobus}), on the contrary, referred to existing NTN
methods as \textit{metaphors} and split them into the following
three groups:
\begin{enumerate}
\item ``spell checking'' metaphor;
\item ``translation'' metaphor;
\item ``speech recognition'' metaphor.
\end{enumerate}

Though both divisions seem to be justified to some extent, it is
difficult to determine using them whether an NTN approach that detects
and restores incorrectly spelled words on the basis of phonetical
\textit{n}-gram statistics should fall into \textit{n}-gram based,
probabilistic, spell checking or speech recognition class.

We instead suggest using two separate types of criteria which are
unrelated to each other and characterize NTN approaches from different
points of view.

The first criterion is \emph{segmentation level}. It depends primarily
on the maximal length of syntagmatic segments which are used
\begin{inparaenum}[\itshape a\upshape)]
\item to infer in-vocabulary (IV) equivalents for OOV
  tokens\label{candidate:inferring} and
\item to choose the most probable variant\label{candidate:selection}
  among multiple possible suggestions\footnote{For the cases, when
    segments of different lengths are used for tasks
    \ref{candidate:inferring} and \ref{candidate:selection}, we note
    it explicitly in our classification on which segmentation length
    each subtask relies.}.
\end{inparaenum} For this criterion we propose division into following
classes:
\begin{enumerate}
  \item graphematic\footnote{Depending on whether phonetical
    information is involved or not at this level, this class could be
    further divided into a phonographematic and purely graphematic
    subclasses.};
  \item lexical;
  \item phrasal.
\end{enumerate}
Each broader level of this hierarchy is also supposed to either
incorporate or ignore information provided by its narrower
subsegments. In this way, we only need to mention one (the broadest)
hierarchical class for the cases when multiple segmentation levels are
involved by some techniques.

The second classification criterion regards the \emph{type of
  information induction} that was used to devise the correction
directives. This leads us to the usual NLP-taxonomy which divides all
approaches into:
\begin{enumerate}
  \item rule-based;
  \item statistical\footnote{Depending on the type of training data
    required, this class is usually divided into unsupervised,
    semi-supervised, and supervised groups.};
  \item and hybrid ones.
\end{enumerate}

Equipped with these two criteria, we could now proceed to the
description of main scientific works on NTN task which were done in
recent years.

It is worthwile noting that many approached to NTN, especially the
earlier ones, were heavily influenced by the idea of noisy channel
model. The underlying notion of this model was that observed sentences
had to be considered as corrupted versions of some original language
signals. In order to restore the undistorted signal, one had to come
up with two probabilistic submodels: the error model (one which caused
the corruption) and the language model (one which reflected the nature
of initial signal source). To these works belong Brill and Moore
(\cite{brill}), Sproat et al. (\cite{sproat}), Toutanova and Moore
(\cite{toutanova}), Clark (\cite{alexander-clark}), Choudhury et
al. (\cite{choudhury}), Cook and Stevenson (\cite{cook}), Beaufort et
al. (\cite{beaufort}) etc. The majority of these methods used either
purely graphematic (Brill and Moore, \cite{brill}; Sproat et al.,
\cite{sproat}; Clark, \cite{alexander-clark}) or phonographematic
segmentation (Toutanova and Moore \cite{toutanova}; Choudhury et al.,
\cite{choudhury}; Cook and Stevenson, \cite{cook}) for error
models. For language model, normally phrasal segments (word
\textit{n}-grams) were used. From the point of view of information
induction almost all of these techniques could be characterized as
supervised statistical, with the exception of Cook and Stevenson
(\cite{cook}) who claimed to use an unsupervised technique.

@@

In our next section will also perform quantitive and qualitative
analyses of unknown words in German Twitter texts in order to assess
whether and to what extent the problem of ill-formed words is relevant
for them. Relying on these analyses we then will suggest a set of
ad-hoc techniques which could help one lower the average rate of
incorrectly spelled words in this kind of text.

@@

Here once again the question of classification arose but this time
with regard to the reasons, why different tokens could have been
omitted from corresponding applications' dictionaries. In this
respect, division into following groups seemed to be appropriate for
us:
\begin{enumerate}
  \item \textbf{Objective limitedness of machine-readable dictionary
    (MRD)}. To this group we counted words of basic vocabulary which
    did not get into applications' MRD either because they supposedly
    were rare or because they did not exist at the time when
    dictionaries were created. Another reason for inclusion in this
    type was the belonging of a word to an open lexical or
    part-of-speech class (like, for example, named entities or
    interjections) which are usually deliberately not included in MRDs
    due to impossibility to cover these classes fully;\label{dict}
  \item \textbf{Unintended sloppiness of user's input}. In the scope
    of this group we considered unintended typos including unexpected
    truncation of words at the end of Twitter messages\footnote{As is
      generally known, Twitter imposes a strong restriction on the
      length of posted messages which can be no longer than 140
      characters. Upon exceeding this length, tweets get automatically
      truncated.};\label{spell}
  \item \textbf{Stylistic specifics of text genre}. This group
    comprised words which could be considered as illegal from the
    point of view of standard language texts but were perfectly valid
    terms in the domain of web discourse and Twitter
    communication.\label{style}
\end{enumerate}

@@

We subdivided group \ref{dict} into the following subgroups:
\begin{enumerate}
\item regular German words, e.g. \textit{aufm}, \textit{losziehen};\label{regular}
\item compounds, e.g.  \textit{Altwein}, \textit{Amtsapothekerin};\label{compound}
\item abbreviations, e.g. \textit{NBG}, \textit{OL};\label{abbr}
\item interjections, e.g.  \textit{aja}, \textit{haha};\label{inj}
\item named entities, with subclasses:\label{ne}
  \begin{enumerate}
  \item persons, e.g.  \textit{Ahmadinedschad}, \textit{Schweiger};
  \item geographic locations, e.g.  \textit{Biel}, \textit{Limmat};
  \item companies, e.g. \textit{Apple}, \textit{Facebook};
  \item product names, e.g. \textit{iPhone}, \textit{MacBook};
  \end{enumerate}
\item neologisms, with subclasses:\label{neolog}
  \begin{enumerate}
    \item newly coined German terms, e.g. \textit{entfolgen},
      \textit{geskypt};\label{new}
    \item loanwords, e.g. \textit{Community},
      \textit{Stream};\label{loan}
  \end{enumerate}
\item and, finally, foreign words like \textit{is} or \textit{now}
  which in contrast to \ref{loan} were not mentioned in any existing
  German lexicons nor they complied with inflectional rules of German
  grammar.\label{fw}
\end{enumerate}
Though this division is admittedly arbitrarily to a certain degree and
involves different linguistic criteria simultaneously, the underlying
notion for it was simple. Valid words could have been omitted from an
MRD either due to the limitation of developers' capacities (group
\ref{regular}), active word formation or lexical productivity of the
language (groups \ref{compound} through \ref{new}) or also due to
language's openness to foreign language systems (groups \ref{loan} and
\ref{fw}).

@@

Similarly to class \ref{dict} we subdivided the group of misspellings
into the following subgroups:
\begin{enumerate}
  \item insertion;
  \item deletion;
  \item substitution;
\end{enumerate}
according to the kind of operation which led to a particular spelling
mistake. The statistics on distribution of these subgroups is shown in
Table \ref{table:spell}.
\begin{table}
  \caption{Distribution of OOV words belonging to the class
    ``unintended sloppiness of user's input''\label{table:spell}}
  \begin{tabular}{p{0.4\textwidth}*{4}{>{\centering\arraybackslash}p{0.15\textwidth}}}
    \hline\noalign{\smallskip}
    \multirow{2}{*}{OOV subclass} & %
    \multicolumn{2}{c}{\texttt{hunspell}} & %
    \multicolumn{2}{c}{\texttt{TreeTagger}}\\
    & \totaloov{} & \uniqoov{} & \totaloov{} & \uniqoov{}\\
    \noalign{\smallskip} \hline
    insertion &  &  &  & \\
    deletion &  &  &  & \\
    substitution &  &  &  & \\\hline
    {\bfseries total} &  &  &  & \\
    \noalign{\smallskip} \hline
  \end{tabular}
\end{table}

>>>>

  This article gives an overview of existing approaches to the problem
  of out-of-vocabulary (OOV) tokens and noisiness phenomena in natural
  language texts. These approaches are classified with regard to the
  size of text spans and knowledge inference mechanisms which they
  rely on in their work. Subsequently, quantitative and qualitative
  analyses of unknown words are conducted on German Twitter messages,
  in order to show how relevant the OOV and text normalization
  problems are for this particular kind of micro-texts and what their
  characteristics are there. In a concluding step, we present a set of
  ad-hoc techniques which are supposed to tackle some of the most
  prominent disturbing effects found during the analyses and show how
  this set of techniques helps us to lower the average rate of
  out-of-vocabulary tokens in Twitter messages thereby improving the
  quality of their automatic Part-of-speech tagging.

@@

Though the problem of out-of-vocabulary words and its closely related
task of textual normalization have been extensively studied in
computational linguistics since as early as the late 1950s
(cf. Petersen, \cite{petersen}) and were certainly anything but new at
the time when mobile communication emerged, it were small messages
that revived interest in this research area in the past two decades.

Starting from the second half of the 1990s, not only the number of
publications on automatic text correction gradually increased every
year but also the focus of those researches steadily turned away from
normalization of manually typed or OCR processed official documents to
the processing of sloppily created user messages and online chat
posts. But even despite this increased interest, most of the
researches still were dealing with only English text data. A few
exceptions from that are works on French (Beaufort et al.,
\cite{beaufort}) and Spanish (Oliva et al., \cite{oliva}).

%% One of the first scientific milestones which marked the renaissance
%% of OOV studies in CL was a comprehensive article by Karen Kukich
%% published in the renowned ACM journal on December 4-th 1992
%% \cite{kukich}. By an interesting coincidence, just exactly one day
%% before that, a British engineer called Neil Papworth had sent the
%% world's first ever SMS-message \cite{guardian:sms}. But in contrast
%% to Papworth's SMS which only said ``Merry Christmas'' to one of his
%% friends, Kukich's work comprised more than 60 pages and provided a
%% fully-fledged review of the state of the art techniques for dealing
%% with unknown words within the scope of spelling correction
%% programs.

In this article we are going to provide some more insight about the
relevance of OOV issues for German language by showing how many
out-of-vocabulary tokens are found in German online texts on average
and what kind of linguistic phenomena are main sources of these
OOVs. The next Section will first give a short overview and provide a
classification of recent scientific approaches to the problem of
out-of-vocabulary words in non-standard texts. After that, in Section
\ref{error:analysis} we will analyze which types of noisiness
phenomena are especially common in German Twitter. Section
\ref{normalization} will subsequently describe an automatic procedure
for mitigating some of the most prominent of those effects. In a
concluding step, we will perform an evaluation of the results of this
procedure and give some further suggestions for future research.

@@

Before proceeding with the description of existing methods for noisy
text normalization (NTN), we first would like to define the criteria
by which these methods could be classified. It should be noted that
there already exist several classifications of NTN approaches
including, for example, Kukich (\cite{kukich}), Kobus et
al. (\cite{kobus}), and Sproat et al. (\cite{sproat}). But all of
these classifications seem to share one significant common
shortcoming, namely they try to involve several independent criteria
within one classification scheme.

@@

Kobus (\cite{kobus}), on the contrary, referred to NTN methods as
\textit{metaphors} and split them into the following three groups:
\begin{enumerate}
\item ``spell checking'' metaphor;
\item ``translation'' metaphor;
\item ``speech recognition'' metaphor.
\end{enumerate}

Though both divisions seem to be justified to some extent, it is
difficult to determine using them whether an NTN approach that detects
and restores incorrectly spelled words on the basis of phonetical
\textit{n}-gram statistics should fall into the \textit{n}-gram based,
probabilistic, spell checking or speech recognition class.

Instead of that, we suggest using two separate types of criteria which
are unrelated to each other and characterize NTN approaches from
different points of view.

The first criterion is \emph{segmentation level}. It primarily depends
on the maximal length of syntagmatic segments which are used
\begin{inparaenum}[\itshape a\upshape)]
\item to infer in-vocabulary (IV) equivalents for OOV
  tokens\label{candidate:inferring} and
\item to choose the most probable variant\label{candidate:selection}
  among multiple possible suggestions\footnote{For the cases, when
    segments of different lengths are used for tasks
    \ref{candidate:inferring} and \ref{candidate:selection}, we note
    it explicitly in our classification on which segmentation length
    each of these subtasks relies.}.
\end{inparaenum} For this criterion we propose division into the following
classes:
\begin{enumerate}
  \item graphematic\footnote{Depending on whether phonetical
    information is involved or not at this level, this class could be
    further divided into a phonographematic and purely graphematic
    subclasses.};
  \item lexical;
  \item phrasal.
\end{enumerate}
Each broader level of this hierarchy is also supposed to either
incorporate or ignore information provided by its narrower
subsegments. In this way, only one (the broadest) hierarchical class
needs to be mentioned for the cases when multiple segmentation levels
are involved by some techniques.

The second classification criterion regards the \emph{type of
  information induction} that is used to devise the correction
directives. This leads us to the usual NLP-taxonomy which divides all
approaches into:
\begin{enumerate}
  \item rule-based;
  \item statistical\footnote{Depending on the type of training data
    required, this class is in turn usually divided into unsupervised,
    semi-supervised, and supervised groups.};
  \item and hybrid ones.
\end{enumerate}

Equipped with these two criteria, it should now be easier to get a
clear picture of the development and recent tendencies in NTN
research.

It is worthwile noting that many approaches to NTN, especially the
earlier ones, were heavily influenced by the idea of noisy channel
model. The underlying notion of this model was that observed sentences
had to be considered as corrupted versions of their respective
original language signals. In order to restore the undistorted signal,
one had to come up with two probabilistic submodels: the error model
(one which caused the corruption) and the language model (one which
reflected the nature of initial signal source). To these works belong
Brill and Moore (\cite{brill}), Sproat et al. (\cite{sproat}),
Toutanova and Moore (\cite{toutanova}), Clark
(\cite{alexander-clark}), Choudhury et al. (\cite{choudhury}), Cook
and Stevenson (\cite{cook}), Beaufort et al. (\cite{beaufort})
etc. The majority of these methods used either purely graphematic
(Brill and Moore, \cite{brill}; Sproat et al., \cite{sproat}; Clark,
\cite{alexander-clark}) or phonographematic segmentation (Toutanova
and Moore \cite{toutanova}; Choudhury et al., \cite{choudhury}; Cook
and Stevenson, \cite{cook}) for error models. For language model,
normally phrasal segments (word \textit{n}-grams) were used. From the
point of view of information induction almost all of these techniques
could be characterized as supervised statistical, with the exception
of Cook and Stevenson (\cite{cook}) who claimed to use an unsupervised
technique.

@@

In the next Section will also perform quantitive and qualitative
analyses of unknown words in German Twitter texts in order to assess
whether and to what extent the problem of ill-formed words is relevant
for them. Relying on these analyses we then will suggest a set of
ad-hoc techniques which could help one lower the average rate of
incorrectly spelled words in this kind of text.

@@

Here once again the question of classification arose but this time
with regard to the reasons, why different tokens could have been
omitted from corresponding applications' dictionaries. In this
respect, division into following groups seemed to be appropriate for
us:
\begin{enumerate}
  \item \textbf{Objective limitedness of machine-readable dictionary
    (MRD)}. To this group we counted words of basic vocabulary which
    did not get into applications' MRD either because they supposedly
    were rare or because they did not exist at the time when
    dictionaries were created. Another reason for inclusion in this
    type was the belonging of a word to an open lexical or
    part-of-speech class (like, for example, named entities or
    interjections) which are usually deliberately not included in MRDs
    due to impossibility to fully cover these classes;\label{dict}
  \item \textbf{Unintended sloppiness of users' input}. In the scope
    of this group we considered unintended typos including unexpected
    truncation of words at the end of Twitter messages\footnote{As is
      generally known, Twitter imposes a strong restriction on the
      length of posted messages which can be no longer than 140
      characters. Upon exceeding this length, tweets get automatically
      truncated to maximal allowed length.};\label{spell}
  \item \textbf{Stylistic specifics of text genre}. This group
    comprised words which could be considered as illegal from the
    point of view of standard language texts but were perfectly valid
    terms in the domain of web discourse and Twitter
    communication.\label{style}
\end{enumerate}

@@

We subdivided group \ref{dict} into the following subgroups:
\begin{enumerate}
\item regular German words, e.g. \textit{aufm}, \textit{losziehen};\label{regular}
\item compounds, e.g.  \textit{Altwein}, \textit{Amtsapothekerin};\label{compound}
\item abbreviations, e.g. \textit{NBG}, \textit{OL};\label{abbr}
\item interjections, e.g.  \textit{aja}, \textit{haha};\label{inj}
\item named entities, with subclasses:\label{ne}
  \begin{enumerate}
  \item persons, e.g.  \textit{Ahmadinedschad}, \textit{Schweiger};
  \item geographic locations, e.g.  \textit{Biel}, \textit{Limmat};
  \item companies, e.g. \textit{Apple}, \textit{Facebook};
  \item product names, e.g. \textit{iPhone}, \textit{MacBook};
  \end{enumerate}
\item neologisms, with subclasses:\label{neolog}
  \begin{enumerate}
    \item newly coined German terms, e.g. \textit{entfolgen},
      \textit{geskypt};\label{new}
    \item loanwords, e.g. \textit{Community},
      \textit{Stream};\label{loan}
  \end{enumerate}
\item and, finally, foreign words like \textit{is} or \textit{now}
  which in contrast to \ref{loan} were not mentioned in any existing
  German lexicons and did not comply with inflectional rules of German
  grammar.\label{fw}
\end{enumerate}
Though this division is admittedly arbitrarily to a certain degree and
involves different linguistic criteria simultaneously, the underlying
notion for it was simple. Valid words could have been omitted from an
MRD either due to the limitation of developers' capacities (group
\ref{regular}), active word formation or lexical productivity of the
language (groups \ref{compound} through \ref{new}) or also due to
language's openness to foreign language systems (groups \ref{loan} and
\ref{fw}).

@@

Similarly to class \ref{dict} we subdivided the group of misspellings
into the following subgroups:
\begin{enumerate}
  \item insertion;
  \item deletion;
  \item substitution;
\end{enumerate}
according to the kind of operation which led to a particular spelling
mistake. The statistics on distribution of these subgroups is shown in
Table \ref{table:spell}.
\begin{table}
  \caption{Distribution of OOV words belonging to the class
    ``unintended sloppiness of users' input''\label{table:spell}}
  \begin{tabular}{p{0.4\textwidth}*{4}{>{\centering\arraybackslash}p{0.15\textwidth}}}
    \hline\noalign{\smallskip}
    \multirow{2}{*}{OOV subclass} & %
    \multicolumn{2}{c}{\texttt{hunspell}} & %
    \multicolumn{2}{c}{\texttt{TreeTagger}}\\
    & \totaloov{} & \uniqoov{} & \totaloov{} & \uniqoov{}\\
    \noalign{\smallskip} \hline
    insertion &  &  &  & \\
    deletion &  &  &  & \\
    substitution &  &  &  & \\\hline
    {\bfseries total} &  &  &  & \\
    \noalign{\smallskip} \hline
  \end{tabular}
\end{table}
